---
title: "R Notebook"
output: html_notebook
---

# Load Libraries

```{r, warnings=F, message=F}
library(Seurat)
library(WGCNA)
library(cluster)
library(parallelDist)
library(ggsci)
library(emmeans)
library(lme4)
library(ggbeeswarm)
library(genefilter)
library(tidyverse)
library(reshape2)
library(igraph)
library(gProfileR)
library(ggpubr)
library(princurve)
library(here)
library(cowplot)
library(tidygraph)
library(ggraph)
```

```{r include = FALSE}
knitr::opts_chunk$set(message = FALSE, warnings = FALSE)
source(here("code/sc_functions.R"))
```

# Extract Cells for WGCNA
```{r, warnings=FALSE, message=FALSE, include=FALSE}
fgf.glia.wc <- readRDS("/projects/dylan/wc_fgf/data/tany_labeled.RDS")
fgf.glia.wc %>%
  SCTransform(return.only.var.genes = F, verbose = F, variable.features.n = 5000) -> ventric
```

# Calculate Pseudoventricle Scores
```{r, fig.height=4, fig.width=3}
pcembed <- as.matrix(Embeddings(ventric, reduction = "pca")[,c(1:2)])
y <- principal_curve(pcembed)
# color <- as.factor(ventric$predicted.id)
# levels(color) <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3")
df = data.frame(y$s[order(y$lambda), ])
colnames(df) = c("x", "y")
points <- data.frame(id = ventric$predicted.id, Embeddings(ventric, reduction = "pca")[,c(1:2)])
rand <- sample(nrow(points), 3000)
princ_plot <- ggplot(data = df, aes(x, y)) +
  geom_point(data = points[rand,], aes(x=PC_1, y=PC_2, colour=factor(id)), size = 2, inherit.aes = F, alpha=0.5) + 
  geom_line(arrow = arrow(length=unit(0.30,"cm"), ends="last", type = "closed"), size = 2, color = "orange") + 
  ggpubr::theme_pubr(legend = "none") + 
  ggsci::scale_color_jco() +
  xlim(c(-20,10)) + coord_flip() + xlab("PC1") + ylab("PC2") +  theme_figure
ventric$height <- y$lambda
princ_plot
```

# Calculate softpower
```{r}
enableWGCNAThreads()
datExpr<-as.matrix(t(ventric[["SCT"]]@scale.data[ventric[["SCT"]]@var.features,]))
gsg = goodSamplesGenes(datExpr, verbose = 3)
powers = c(c(1:10), seq(from = 12, to=40, by=2))
sft=pickSoftThreshold(datExpr,dataIsExpr = TRUE, powerVector = powers, corOptions = list(use = 'p'), 
                      networkType = "signed")
cex1=0.9
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit, signed R^2",type="n", main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],labels=powers ,cex=cex1,col="red")
abline(h=0.80,col="red")
#Mean Connectivity Plot
plot(sft$fitIndices[,1], sft$fitIndices[,5],xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

# Generate TOM
```{r}
softPower <- 12
SubGeneNames <- colnames(datExpr)
adj <- adjacency(datExpr, type = "signed", power = softPower)
diag(adj) <- 0
TOM <- TOMsimilarityFromExpr(datExpr, networkType = "signed", TOMType = "signed", power = softPower, maxPOutliers = 0.05)
colnames(TOM) <- rownames(TOM) <- SubGeneNames
dissTOM <- 1 - TOM
geneTree <- hclust(as.dist(dissTOM), method = "complete") # use complete for method rather than average (gives better results)
plot(geneTree, xlab = "", sub = "", cex = .5, main = "Gene clustering", hang = .001)
```

# Identify Modules
```{r Cut Gene Tree}
minModuleSize = 15
x = 2
dynamicMods = cutreeDynamic(dendro = geneTree, distM = as.matrix(dissTOM), 
                            method="hybrid", pamStage = F, deepSplit = x, 
                            minClusterSize = minModuleSize)
dynamicColors = labels2colors(dynamicMods) #label each module with a unique color
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05, 
                    main = "Gene dendrogram and module colors") #plot the modules with colors
```

# Calculate Eigengenes and Merge Close Modules
```{r merge close modules}
MEs = moduleEigengenes(datExpr, dynamicColors)$eigengenes #this matrix gives correlations between cells and module eigengenes (a high value indicates that the cell is highly correlated with the genes in that module)
ME1<-MEs
row.names(ME1)<-row.names(datExpr)
# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);
# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average");
# Plot the result
plot(METree, main = "Clustering of module eigengenes",xlab = "", sub = "")
MEDissThres = 0.2
# Plot the cut line into the dendrogram
abline(h=MEDissThres, col = "red")
```

# The merged module colors
```{r}
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
mergedColors = merge$colors
mergedMEs = merge$newMEs
moduleColors = mergedColors
MEs = mergedMEs
modulekME = signedKME(datExpr,MEs)
```

# Plot merged modules
```{r}
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
c("Dynamic Tree Cut", "Merged dynamic"),
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
# colorOrder = c("grey", standardColors(50));
# moduleLabels = match(moduleColors, colorOrder)-1
MEs = mergedMEs
modulekME = signedKME(datExpr,MEs)
```

# Type gene name, prints out gene names also in that module
```{r}
modules<-MEs
c_modules<-data.frame(moduleColors)
row.names(c_modules)<-colnames(datExpr) #assign gene names as row names
module.list.set1<-substring(colnames(modules),3) #removes ME from start of module names
index.set1<-0
Network=list() #create lists of genes for each module
for (i in 1:length(module.list.set1)){index.set1<-which(c_modules==module.list.set1[i])
Network[[i]]<-row.names(c_modules)[index.set1]}
names(Network)<-module.list.set1
lookup<-function(gene,network){return(network[names(network)[grep(gene,network)]])} #load function
```

# Filter metadata table and correlate with eigengenes
```{r, fig.height=6, fig.width=5}
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
MEs = orderMEs(MEs)
MEs %>% dplyr::select(-MEgrey) -> MEs
var<-model.matrix(~0+ventric$trt)
moduleTraitCor <- cor(MEs, var, use="p")
cor<-moduleTraitCor[abs(moduleTraitCor[,1])>.2,]
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
cor<-melt(cor)
ggplot(cor, aes(Var2, Var1)) + geom_tile(aes(fill = value), 
     colour = "white") + scale_fill_gradient2(midpoint = 0, low = "blue", mid = "white",
                            high = "red", space = "Lab", name="Correlation \nStrength") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Treatment") + ylab(NULL)
```
# Get hubgenes
```{r}
hubgenes<-lapply(seq_len(length(Network)), function(x) {
  dat<-modulekME[Network[[x]],]
  dat<-dat[order(-dat[paste0("kME",names(Network)[x])]),]
  gene<-rownames(dat)
  return(gene)
})

names(hubgenes)<-names(Network)
d <- unlist(hubgenes)
d <- data.frame(gene = d, 
           vec = names(d))

write_csv(d, path=here("output/glia/wgcna/wc_tany_wgcna_genemodules.csv"))
```

# Linear regression
```{r, fig.height=20, fig.width=20}
data<-data.frame(MEs, trt = ventric$trt, 
                 sample = as.factor(ventric$sample),
                 batch = as.factor(ventric$batch), 
                 height = ventric$height,
                 bin = cut(ventric$height, seq_len(max(ventric$height))),
                 type = ventric$predicted.id)

levels(data$bin) <- c(1:74)
data$bin <- as.numeric(as.character(data$bin))
data %>% filter(bin>=12 & bin <= 40) %>% mutate(bin=factor(bin)) -> data

plot <- lapply(colnames(MEs), function(x) {
  x <- data.frame(scale(data[,x]))
  x$bin<-data$bin
  x$trt<-as.factor(data$trt)
  x$type<-as.factor(data$type)
  x<-melt(x, id.vars=c("trt","bin","type"))
  x<-x[complete.cases(x),]
  x %>% dplyr::group_by(trt, bin) %>%
    dplyr::summarise(mean = mean(value), sd=sd(value), se = sd/sqrt(length(value))) ->plotval
  return(plotval)
})

names(plot) <- colnames(MEs)
plot_df <- bind_rows(plot, .id="id")
```

# Plotting
```{r, fig.height=12, fig.width=12}
mod <- lapply(colnames(MEs)[grepl("^ME", colnames(MEs))], function(me) {
  tryCatch({
    mod <- lmer(data[[me]] ~ trt*bin + (1 | batch) + (1 | sample), data = data)
    pairwise <- emmeans(mod, pairwise ~ trt|bin)
    plot <- data.frame(plot(pairwise, plotIt = F)$data)
    sig <- as.data.frame(pairwise$contrasts)
    return(sig) }, error = function(err) {
    print(err)
  }
  )
})

names(mod) <- colnames(MEs)
mod_df <- bind_rows(mod, .id="id")
mod_df$p.adj<-p.adjust(mod_df$p.value)
plot_df %>% filter(trt=="FGF") %>% mutate(p.adj = mod_df$p.adj) -> plot_f_df
plot_df %>% filter(trt=="PF") %>% mutate(p.adj = mod_df$p.adj) -> plot_p_df
plot_df <- rbind(plot_f_df, plot_p_df)
plot_df%>%mutate(signif=ifelse(p.adj>.05, "ns",
                                     ifelse(p.adj<.05&p.adj>.01, "*",
                                            ifelse(p.adj<.01&p.adj>.001, "**",
                                                   "***")))) -> plotval_frame

plotval_frame$signif[plotval_frame$signif=="ns"]<-NA
plotval_frame$signif[plotval_frame$trt!="FGF"]<-NA

detach("package:here", unload = T)
library(here)
write_csv(as.data.frame(plotval_frame), path = here("output/glia/wgcna/wc_tany_pseudovent_linmod.csv"))

mod_df %>% dplyr::group_by(id) %>% 
  dplyr::summarise(order = quantile(-log10(p.adj),.75)*(quantile(abs(estimate),.75))) %>% 
  arrange(-order) %>% 
  mutate(pubname = paste0("sc-tany-M",seq_len(length(order)))) %>% dplyr::select(1,3) -> rename
write_csv(as.data.frame(rename), path = here("output/glia/wgcna/wc_tany_colortopubname.csv"))

mod_df %>% filter(p.adj<0.05, estimate>0) -> sig_df
sig_mods <- names(which(table(sig_df$id) > 5))
height_type <- ggplot() + geom_density(data=data, aes(x=(-height), fill=type), inherit.aes = F, alpha=0.25) + coord_flip() + 
  theme_pubr(legend = "none") + xlab("Pseudo-ventricle Height") + ylab(NULL) + scale_fill_jco() + theme(axis.text.x = element_blank(),  axis.text.y = element_blank()) + theme_figure
plotval_frame$pubname <- rename$pubname[match(plotval_frame$id, rename$id)]

mod <- ggplot(plotval_frame, 
              aes(x=(-as.numeric(as.character(bin))), y=mean, color=trt, label=signif)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, width=.1)) +
  geom_line() + geom_point() + scale_color_manual(values=c("#000000","#999999")) +
  geom_text(color="black",size=3,aes(y= mean + .5), position=position_dodge(.9), angle=90) +
  coord_flip() +
  facet_wrap(vars(pubname), scales="free", nrow=4) + theme_pubr() + ylab("Scaled ME Expression") +
  xlab(NULL) +
  theme(legend.position="none", axis.text.y = element_blank(), 
        axis.text.x = element_text(angle=45, hjust=1)) + theme_figure
mod
ggsave(here("data/figures/supp/alltany_mods_pseudovent.pdf"), width = 12, h=12)

mod_sig <- ggplot(plotval_frame[plotval_frame$id%in%c(sig_mods),], aes(x=(-as.numeric(as.character(bin))), y=mean, color=trt, label=signif)) + 
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, width=.1)) +
  geom_line() + geom_point() + scale_color_manual(values=c("#000000","#999999")) +
  geom_text(color="black",size=3,aes(y= mean + .5), position=position_dodge(.9), angle=90) + coord_flip() +
  facet_wrap(vars(pubname), scales="free", nrow=1) + theme_pubr() + ylab("Scaled ME Expression") + xlab(NULL) +
  theme(legend.position="none", axis.text.y = element_blank(), axis.text.x = element_text(angle=45, hjust=1)) + theme_figure
```
# UMAP plot
```{r}
ventric_umap <- as.data.frame(Embeddings(ventric, reduction="umap"))
ventric_umap$`Cell Type` <- ventric$predicted.id
ggplot(ventric_umap, aes(x=UMAP_1, y=UMAP_2, color=`Cell Type`)) + 
  geom_point(alpha=0.5) + xlab("UMAP1") + ylab("UMAP2") + ggpubr::theme_pubr() + 
  ggsci::scale_color_jco(labels=c(expression(alpha[1]~Tanycyte), expression(alpha[2]~Tanycyte),
                                  expression(beta[1]~Tanycyte), expression(beta[2]~Tanycyte))) +
  coord_cartesian(clip="off") +
  guides(color = guide_legend(override.aes = list(size=3))) +
  theme(legend.position = c(0.93,0.2), legend.background = element_blank(), 
        legend.title = element_blank(), legend.key.size = unit(0, 'lines')) -> umap
umap
```

# Arrange plot
```{r, fig.height=4, fig.width=12}
tany_day5 <- plot_grid(umap, princ_plot, height_type, ggplot() + theme_void(), mod_sig, rel_widths = c(2,1,1,1, 2), align = "hv", axis="tb", nrow=1, scale=0.9, labels=c("a","b","","","c"))
tany_day5
```
# Calculate GO term enrichment
```{r}
goterms<-lapply(hubgenes[gsub(sig_mods,pattern = "ME",replacement = "")], function(x) { 
  x<-gprofiler(x, ordered_query = T, organism = "mmusculus", significant = T, custom_bg = colnames(datExpr),
                           src_filter = c("GO:BP","REAC","KEGG"), hier_filtering = "strong",
                           min_isect_size = 2, 
                           sort_by_structure = T,exclude_iea = T, 
                           min_set_size = 10, max_set_size = 300,correction_method = "fdr")
  x<-x[order(x$p.value),]
  return(x)
})

goterms %>% bind_rows(.id="id") %>%
  mutate(padj=p.adjust(p.value, "fdr")) -> godat

write_csv(godat, path=here("output/glia/wgcna/wc_wgcna_tany_goterms.csv"))
save.image(file = here("output/glia/wgcna/wc_tany_results.RData"))
```
# Plot GO terms
```{r}
goterms[[1]] %>% 
  select(domain, term.name, p.value, overlap.size) %>% arrange(p.value) %>% top_n(5, -p.value) %>%
  mutate(x = fct_reorder(str_to_title(str_wrap(term.name,20)), -p.value)) %>% 
  mutate(y = -log10(p.value)) %>%
  ggplot(aes(x,y)) + 
  geom_col(colour="black", width = 1, fill="gray80", size=1) +
  theme_pubr(legend="none") + 
  theme(axis.text.y = element_text(size=8)) +
  scale_size(range = c(5,10)) + 
  ggsci::scale_fill_lancet() +
  coord_flip() +
  xlab(NULL) + ylab(expression(bold(-log[10]~pvalue))) +
  theme_figure -> tany_sc1
tany_sc1

goterms[[3]] %>% 
  select(domain, term.name, p.value, overlap.size) %>% arrange(p.value) %>% top_n(5, -p.value) %>%
  mutate(x = fct_reorder(str_to_title(str_wrap(term.name,20)), -p.value)) %>% 
  mutate(y = -log10(p.value)) %>%
  ggplot(aes(x,y)) + 
  geom_col(colour="black", width = 1, fill="gray80", size=1) +
  theme_pubr(legend="none") + 
  theme(axis.text.y = element_text(size=8)) +
  scale_size(range = c(5,10)) + 
  ggsci::scale_fill_lancet() +
  coord_flip() +
  xlab(NULL) + ylab(expression(bold(-log[10]~pvalue))) +
  theme_figure -> tany_sc3
tany_sc3
```
# Gene network plots
```{r}
hubgenes <- lapply(seq_len(length(Network)), function(x) {
  dat <- modulekME[Network[[x]], ]
  dat <- dat[order(-dat[paste0("kME", names(Network)[x])]), ]
  gene <- data.frame(gene=rownames(dat),kme=dat[,x]) 
  return(gene)
})


names(hubgenes) <- names(Network)

color <- c("darkgreen")
lapply(color, function(col) {
  maxsize <- 15
  hubs <- data.frame(genes=hubgenes[[col]]$gene[1:maxsize], kme = hubgenes[[col]]$kme[1:maxsize], mod = rep(col,15))
}) -> hub_plot 


hub_plot <- lapply(hub_plot, function(x) {
  adj[as.character(x$genes), as.character(x$genes)] %>% 
  graph.adjacency(mode = "undirected", weighted = T, diag = FALSE) %>%
  as_tbl_graph(g1) %>% upgrade_graph() %>% activate(nodes) %>% dplyr::mutate(mod=x$mod) %>%
  dplyr::mutate(kme=x$kme)  %>% 
  activate(edges)}
)

hub_plot[[1]] %>% activate(nodes) %>% mutate(color = ifelse(name == "Vim", yes="red", no = "black")) -> hub_plot

set.seed("139")
ggraph(hub_plot,  layout = 'kk') + 
  geom_edge_link(color="darkgrey", aes(alpha = weight), show.legend = F)  + 
  scale_edge_width(range = c(0.2, .5)) + geom_node_text(aes(label = name, color=color), fontface="bold", size=4, repel=T) +
  geom_node_point(shape=21, alpha=0.5, fill="grey70", size=3) + scale_color_manual(values=c("gray0","red")) +
  theme_graph() + theme(legend.position = "none", plot.title = element_text(hjust=0.5, vjust=1)) + 
    coord_cartesian(clip = "off") -> genenet
genenet
```
# Vimentin quantification
```{r}
readxl::read_xlsx(here("data/mouse_data/fig5/191118_Vim_quantification.xlsx"), range="A5:B9") %>%
  reshape2::melt() -> vim

vim %>% dplyr::group_by(variable) %>% dplyr::summarise(mean = mean(value), sd = sd(value), se = sd/sqrt(length(value))) %>%
  ggplot(aes(x = variable, y = mean, fill = variable, color = variable)) + 
  geom_col(width=1, alpha=0.75, colour="black", position="dodge") +
  geom_errorbar(aes(x=variable, ymin = mean-se, ymax=mean+se), width=0.2, position = position_dodge(0.9)) + 
  geom_jitter(data = vim, inherit.aes = F, aes(x=variable, y=value, fill=variable), 
              alpha=0.5, shape=21, position = position_jitterdodge(.5)) + xlab(NULL) + 
  geom_signif(y_position=c(7.5e6), xmin=c(1.2), xmax=c(1.8),
              annotation=c("*"), tip_length=0, size = 0.5, textsize = 5, color="black") +
  ylab(expression(bold(Volume~(mu*m^3)))) + scale_fill_manual(values=c("gray80","gray30"), name="") + 
  scale_color_manual(values=c("gray80","gray30"), name="") +
  theme_pubr() + theme(legend.position = "none", axis.text.x = element_text(angle=45,hjust=1)) + theme_figure -> vim_quant

vim_plot <- cowplot::plot_grid(ggplot() + theme_void(), vim_quant, nrow=1, scale=0.8, labels=c("e"), rel_widths = c(2,1))
```
# Arrange bottom half of plot
```{r, fig.width=12, fig.height=4}
sc1 <- plot_grid(genenet, tany_sc1, scale=c(1,.8), labels="d")
tany_bot <- plot_grid(sc1, vim_plot, rel_widths = c(1.25,1))
tany_bot
```

# Arrange full figure
```{r fig.width=12, fig.height=7}
plot_grid(tany_day5, tany_bot, ncol=1, rel_heights = c(1.25,1))
ggsave(filename = here("data/figures/fig5/fig5.tiff"), width = 12, h=7, compression="lzw")
```

