---
title: "Glial DGE"
output: html_notebook
---

# Load Libraries
```{r, warning=F, message=F}
library(Seurat)
library(DESeq2)
library(future.apply)
library(cowplot)
library(tidyverse)
library(ggrepel)
library(reshape2)
library(ggpubr)
library(here)
library(wesanderson)
library(ggupset)
library(ggcorrplot)
library(gProfileR)
plan(multiprocess, workers=40)
options(future.globals.maxSize = 4000 * 1024^2)
```

```{r include = FALSE}
knitr::opts_chunk$set(message = FALSE, warnings = FALSE)
source(here("code/sc_functions.R"))
```

# Generate Glial Plots
```{r}
fgf.glia.sub<-readRDS(here("data/glia/glia_seur_filtered.RDS"))
fgf.glia.sub <- RenameIdents(fgf.glia.sub, "COP" = "OPC_COP")

fgf.glia.sub$group<-paste0(fgf.glia.sub$trt, "_", fgf.glia.sub$day)
data.frame(Embeddings(fgf.glia.sub, reduction = "umap")) %>% 
  mutate(group = fgf.glia.sub$group) %>%
  mutate(celltype = Idents(fgf.glia.sub)) %>%
  .[sample(nrow(.)),] %>%
  mutate(group = replace(group, group == "FGF_Day-5", "FGF_d5")) %>% 
  mutate(group = replace(group, group == "FGF_Day-1", "FGF_d1")) %>% 
  mutate(group = replace(group, group == "PF_Day-1", "Veh_d1")) %>%
  mutate(group = replace(group, group == "PF_Day-5", "Veh_d5")) -> umap_embed
colnames(umap_embed)[1:2] <- c("UMAP1", "UMAP2")

label.df <- data.frame(cluster=levels(umap_embed$celltype),label=levels(umap_embed$celltype))
label.df_2 <- umap_embed %>% 
  dplyr::group_by(celltype) %>% 
  dplyr::summarize(x = median(UMAP1), y = median(UMAP2))

p1<-ggplot(umap_embed, aes(x=UMAP1, y=UMAP2, colour=celltype)) + 
  geom_point(alpha=0.5, size=2)  + 
  geom_text_repel(data = label.df_2, aes(label = celltype, x=x, y=y), 
                  size=3, fontface="bold", inherit.aes = F, bg.colour="white")  +
  theme_pubr(legend="none") + ggsci::scale_color_igv() + theme_figure
p1
```
# Integration across timepoints and treatment
```{r}
p2<-ggplot(umap_embed, aes(x=UMAP1, y=UMAP2, colour=group)) + 
  geom_point(alpha=.5, size=1) + 
  ggsci::scale_color_igv() + 
  theme_pubr(legend = "none") + theme_figure
p2
```

# Get colors for matching
```{r}
g <- ggplot_build(p1)
cols<-data.frame(colours = as.character(unique(g$data[[1]]$colour)), 
             label = as.character(unique(g$plot$data[, g$plot$labels$colour])))
colvec<-as.character(cols$colours)
names(colvec)<-as.character(cols$label)
```

# Generate Pseudo Counts
```{r, warning=FALSE, message=FALSE}
fgf.glia.sub<-ScaleData(fgf.glia.sub, verbose=F)
split_mats<-splitbysamp(fgf.glia.sub, split_by="sample")
names(split_mats)<-unique(Idents(fgf.glia.sub))
pb<-replicate(100, gen_pseudo_counts(split_mats, ncells=10)) 
names(pb)<-paste0(rep(names(split_mats)),rep(1:100, each=length(names(split_mats))))
```

# Generate DESeq2 Objects
```{r, warning=FALSE, message=FALSE}
res<-rundeseq(pb)
```

# Identify most responsive cell types
```{r}
degenes<-lapply(res, function(x) {
  tryCatch({
    y<-x[[2]]
    y<-na.omit(y)
    data.frame(y)%>%filter(padj<0.1)%>%nrow()}, 
    error=function(err) {NA})
})

boxplot<-lapply(unique(Idents(fgf.glia.sub)), function(x) {
  y<-paste0("^",x)
  z<-unlist(degenes[grep(y, names(degenes))])
})


names(boxplot)<-unique(Idents(fgf.glia.sub))
genenum<-melt(boxplot)
colnames(genenum)<-c("number","CellType")
genenum <- write_csv(genenum, path = here("output/glia/glia_resampling_output.csv"))
deplot_re <- ggplot(genenum, aes(x=reorder(CellType, -number), y=number, fill=CellType)) + 
  geom_boxplot(notch = T, alpha=1) + scale_fill_manual(values = colvec)  +
  ylab("Number DEG") + xlab(NULL)  + theme_pubr(legend="none") + theme_figure
deplot_re
```

# Generate Pseudo Counts
```{r}
split_mats<-lapply(unique(Idents(fgf.glia.sub)), function(x){
  sub<-subset(fgf.glia.sub, idents=x)
  DefaultAssay(sub)<-"SCT"
  list_sub<-SplitObject(sub, split.by="sample")
  return(list_sub)
})
names(split_mats)<-unique(Idents(fgf.glia.sub))

pseudo_counts<-lapply(split_mats, function(x){
  lapply(x, function(y) {
    DefaultAssay(y) <- "SCT"
    mat<-GetAssayData(y, slot="counts")
    counts <- Matrix::rowSums(mat)
    }) %>% do.call(rbind, .) %>% t() %>% as.data.frame()
})

names(pseudo_counts)<-names(split_mats)
```

# Generate DESeq2 Objects
```{r, warning=FALSE, message=FALSE}
dds_list<-lapply(pseudo_counts, function(x){
  tryCatch({
      trt<-ifelse(grepl("FGF", colnames(x)), yes="F", no="P")
      number<-sapply(strsplit(colnames(x),"_"),"[",1)
      day<-ifelse(as.numeric(as.character(number))>10, yes="5", no="1")
      meta<-data.frame(trt=trt, day=factor(day))
      dds <- DESeqDataSetFromMatrix(countData = x,
                                    colData = meta,
                                    design = ~ 0 + trt)
      dds$group<-factor(paste0(dds$trt, "_", dds$day))
      design(dds) <- ~ 0 + group
      keep <- rowSums(counts(dds) >= 5) > 5
      dds <- dds[keep,]
      dds<-DESeq(dds)
      res_5<-results(dds, contrast = c("group","F_5","P_5"))
      res_1<-results(dds, contrast = c("group","F_1","P_1"))
      f_5_1<-results(dds, contrast = c("group","F_5","F_1"))
      p_5_1<-results(dds, contrast = c("group","P_5","P_1"))
      return(list(dds, res_1, res_5,f_5_1, p_5_1))
    }, error=function(err) {print(err)})
})
```

# Volcano Plot of DE genes
```{r volcano plot, fig.width=12, fig.height=6}
volc_list<-lapply(dds_list, function(x) {
  x[[2]] %>% na.omit() %>% data.frame() %>% add_rownames("gene") %>% 
    mutate(siglog=ifelse(padj<0.05&abs(log2FoldChange)>1, yes=T, no=F)) %>% 
    mutate(onlysig=ifelse(padj<0.05&abs(log2FoldChange)<1, yes=T, no=F)) %>% 
    mutate(onlylog=ifelse(padj>0.05&abs(log2FoldChange)>1, yes=T, no=F)) %>% 
    mutate(col=ifelse(siglog==T, yes="1", no = 
                      ifelse(onlysig==T, yes="2", no = 
                               ifelse(onlylog==T, yes="3", no="4")))) %>% 
    arrange(padj) %>% mutate(label=ifelse(min_rank(padj) < 15, gene, "")) %>% 
    dplyr::select(gene, log2FoldChange, padj, col, label)
})

mapply(x = volc_list, y = names(volc_list), function(x, y) {
  write_csv(x, path = here(sprintf("output/glia/nuclei/d1_%s_pseudobulk_dge.csv",y)))
})

plotlist<-mapply(x=volc_list, y=names(volc_list), function(x,y){
    ggplot(x, aes(y=(-log10(padj)), x=log2FoldChange, colour=factor(col), label=label)) + 
    xlab(expression(Log[2]*~Fold*~Change)) + ylab(expression(-Log[10]*~pvalue)) + 
    geom_point(size=3, alpha=0.75) + geom_hline(yintercept = -log10(0.05), linetype="dashed") + 
    geom_vline(xintercept = c(-1,1), linetype="dashed") + geom_text_repel(colour="black") + theme_pubr() +
    theme(legend.position = "none", title = element_text(vjust=0.5)) + 
    scale_colour_manual(values = wes_palette("Royal1", 3, type="discrete")[c(2,1,3)]) + 
    ggtitle(y)}, SIMPLIFY = FALSE)
plot_grid(plotlist = plotlist, ncol=3)
ggsave(here("data/figures/supp/allvolplots.pdf"))
devolc_plot <- plot_grid(plotlist=plotlist[c("Astro","Tany")], ncol=2)
```
# Go Term enrichment across cell-types
```{r}
pos_genes <- lapply(dds_list[c("Tany","Astro","Epend")], function(x) {
  x[[2]] %>% na.omit() %>% data.frame() %>% add_rownames("gene") %>% 
    filter(padj<0.05, log2FoldChange>1) %>% pull("gene")
})

neg_genes <- lapply(dds_list[c("Tany","Astro","Epend")], function(x) {
  x[[2]] %>% na.omit() %>% data.frame() %>% add_rownames("gene") %>% 
    filter(padj<0.05, log2FoldChange<(-1)) %>% pull("gene")
})

pos_path <- lapply(pos_genes, function(x) {
  gprofiler(x, organism = "mmusculus", significant = T, 
                           src_filter = c("GO:BP","REAC", "KEGG"), hier_filtering = "strong",
                           min_isect_size = 3, 
                           sort_by_structure = T,exclude_iea = T, 
                           min_set_size = 10, max_set_size = 500,correction_method = "fdr") %>%
    arrange(p.value) 
})

neg_path <- lapply(neg_genes, function(x) {
  gprofiler(x, organism = "mmusculus", significant = T, 
                           src_filter = c("GO:BP","REAC", "KEGG"), hier_filtering = "strong",
                           min_isect_size = 3, 
                           sort_by_structure = T,exclude_iea = T, 
                           min_set_size = 10, max_set_size = 500,correction_method = "fdr") %>%
    arrange(p.value)
})

pos_go <- bind_rows(pos_path, .id="id")
write_csv(pos_go, here("output/glia/nucd1_pos_go.csv"))
neg_go <- bind_rows(neg_path, .id="id")
write_csv(neg_go, here("output/glia/nucd1_neg_go.csv"))
```
# GO term plot of upregulated genes
```{r, fig.width=5, fig.height=4}
pos_go %>% dplyr::group_by(id) %>% dplyr::slice(1:5) %>% dplyr::pull(term.id) -> go_id
pos_go %>% filter(term.id%in%go_id)  %>% 
  ggplot(aes(x=fct_relevel(id,"Tany","Epend","Astro"), y=str_to_title(str_wrap(term.name, 40), locale = "en"))) +
  geom_point(aes(size=(-log10(p.value)), fill=domain), shape=21,  alpha=0.5) +
  scale_size(name = expression(bold(-log[10]*pvalue)), range=c(3,8)) + 
  ggsci::scale_fill_npg(name = "Database", labels=c("GO:BP","KEGG","REAC")) + 
  xlab(NULL) + ylab(NULL) + theme_bw() + theme_figure +
  theme(axis.text.y = element_text(size=7, face="bold")) + 
  guides(fill = guide_legend(override.aes = list(size=4), title.theme = element_text(face="bold", size=8), 
                             label.theme = element_text(face="bold", size=8)),
         size =  guide_legend(title.theme = element_text(size=8), 
                              label.theme = element_text(face="bold", size=8)))  -> pos_go_plot
pos_go_plot
```
# Go term plot of downregulated genes
```{r}
neg_go %>% dplyr::group_by(id) %>% dplyr::slice(1:5) %>% ggplot(aes(x=fct_relevel(id,"Tany","Epend","Astro"), y=str_wrap(term.name, 30))) + 
  geom_point(aes(size=(-log10(p.value)), fill=domain), shape=21, alpha=0.75) +
  scale_size(name = expression(log[10]*pvalue), range=c(3,10)) + ggsci::scale_fill_npg(name = "Database") + xlab(NULL) + ylab(NULL) + 
  theme_figure + guides(fill = guide_legend(override.aes = list(size=6))) -> neg_go_plot
neg_go_plot
```

# Overlap between cell-types
```{r}
res_glia_1<-lapply(dds_list, function(x) {
  data.frame(x[[2]]) %>% add_rownames("gene") %>% na.omit(x) %>% 
    filter(padj<0.05) %>% arrange(padj) %>% select(gene) -> x
})

resglia<-bind_rows(res_glia_1, .id="id")
resglia %>%
  dplyr::group_by(gene) %>%
  dplyr::summarize(Celltype = list(id)) -> resglia

upset <- ggplot(resglia, aes(x=Celltype)) +
    geom_bar(fill=c(rep("black",3),"#E64B35B2","#E64B35B2","#E64B35B2", rep("black",4))) + theme_pubr() +
    scale_x_upset(n_intersections = 10) + xlab(NULL) + ylab("Number DEG") + theme_figure
upset
```

# Generate part of figure
```{r, fig.height=7, fig.width=12}
top <- plot_grid(p1, deplot_re, labels=c("a","b"), scale=0.95, align="hv", axis="tb")
mid <- plot_grid(pos_go_plot, upset, axis="t", scale=0.95, align="hv", labels=c("c","d"), rel_widths = c(1,1))
fig3_top <- plot_grid(top, mid, ncol=1, align="hv", axis="tblr", rel_heights = c(1,1.1))
fig3_top
ggsave2(fig3_top, filename = here("data/figures/fig3/fig3_top.png"), h=7, w=12)
save(fig3_top, file = here("data/figures/fig3/fig3_top.RData"))
```

# Correlation (Supplementary figure)
```{r}
library(ggcorrplot)
ranks<-lapply(dds_list, function(x) {
  x<-data.frame(x[[2]])
  x<-na.omit(x)
  y <- (-log10(x$pvalue))*(x$log2FoldChange)
  z <- rownames(x)
  df<-data.frame(order=y,gene=z)
  df<-df[order(-df$order),]
})

corframe<-Reduce(function(x, y) merge(x, y, all=T, by=c("gene")), ranks)
colnames(corframe)<-c("gene",names(ranks))
corframe<-corframe[,-1]
dim(corframe[complete.cases(corframe),])
plotcor <- cor(corframe, method = "spearman", use="complete.obs")
ggcorrplot(plotcor, hc.order = T, type="lower") + 
  ggsci::scale_fill_gsea(limit = c(0,1))
```
