---
title: "Ventricle WGCNA"
output: html_notebook
---

# Load Libraries
```{r, warnings=F, message=F}
library(Seurat)
library(WGCNA)
library(cluster)
library(parallelDist)
library(ggsci)
library(emmeans)
library(lme4)
library(ggbeeswarm)
library(genefilter)
library(tidyverse)
library(reshape2)
library(igraph)
library(gProfileR)
library(ggpubr)
library(here)
library(ggforce)
library(tidygraph)
library(igraph)
library(ggraph)
library(cowplot)
library(future)
plan("multiprocess", workers=40)
options(future.globals.maxSize = 4000 * 1024^2)
```

```{r include = FALSE}
knitr::opts_chunk$set(message = FALSE, warnings = FALSE)
source(here("code/sc_functions.R"))
```

# Extract Cells for WGCNA
```{r, warnings=FALSE, message=FALSE, include=F}
fgf.glia.sub <- readRDS(here("data/glia/glia_seur_filtered.RDS"))
fgf.glia.sub %>% subset(ident=c("Tany", "Epend", "Astro")) %>% 
  SCTransform(return.only.var.genes = F, verbose = F, variable.features.n = 5000) -> ventric
```

# Calculate softpower
```{r}
enableWGCNAThreads()
datExpr<-as.matrix(t(ventric[["SCT"]]@scale.data[ventric[["SCT"]]@var.features,]))
gsg = goodSamplesGenes(datExpr, verbose = 3);
powers = c(c(1:10), seq(from = 12, to=40, by=2))
sft=pickSoftThreshold(datExpr,dataIsExpr = TRUE, powerVector = powers, corOptions = list(use = 'p'), 
                      networkType = "signed")
cex1=0.9
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit, signed R^2",type="n", main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],labels=powers ,cex=cex1,col="red")
abline(h=0.80,col="red")
#Mean Connectivity Plot
plot(sft$fitIndices[,1], sft$fitIndices[,5],xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```

# Generate TOM
```{r}
softPower = 3
SubGeneNames<-colnames(datExpr)
adj= adjacency(datExpr, type = "signed", power = softPower)
diag(adj)<-0
TOM=TOMsimilarityFromExpr(datExpr, networkType = "signed", TOMType = "signed", power = softPower, maxPOutliers = 0.05)
colnames(TOM) = rownames(TOM) = SubGeneNames
dissTOM=1-TOM
geneTree = hclust(as.dist(dissTOM),method="average") #use complete for method rather than average (gives better results)
plot(geneTree,xlab="",sub="",cex=.5,main="Gene clustering",hang=.001)
```

# Identify Modules
```{r Cut Gene Tree}
minModuleSize = 15
x = 4 
dynamicMods = cutreeDynamic(dendro = geneTree, distM = as.matrix(dissTOM), 
                            method="hybrid", pamStage = F, deepSplit = x, 
                            minClusterSize = minModuleSize)
dynamicColors = labels2colors(dynamicMods) #label each module with a unique color
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05, 
                    main = "Gene dendrogram and module colors") #plot the modules with colors
```

# Calculate Eigengenes and Merge Close Modules
```{r merge close modules}
MEs = moduleEigengenes(datExpr, dynamicColors)$eigengenes 
ME1<-MEs
row.names(ME1)<-row.names(datExpr)
MEDiss = 1-cor(MEs);
METree = hclust(as.dist(MEDiss), method = "average");
plot(METree, main = "Clustering of module eigengenes",xlab = "", sub = "")
MEDissThres = 0.2
abline(h=MEDissThres, col = "red")
```

# Merge modules
```{r}
merge = mergeCloseModules(datExpr, dynamicColors, cutHeight = MEDissThres, verbose = 3)
mergedColors = merge$colors
mergedMEs = merge$newMEs
moduleColors = mergedColors
MEs = mergedMEs
modulekME = signedKME(datExpr,MEs)
```

# Plot merged modules
```{r}
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
c("Dynamic Tree Cut", "Merged dynamic"),
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
moduleColors = mergedColors
MEs = mergedMEs
modulekME = signedKME(datExpr,MEs)
```

# Generate function to look up genes in each network
```{r}
modules<-MEs
c_modules<-data.frame(moduleColors)
row.names(c_modules)<-colnames(datExpr) 
module.list.set1<-substring(colnames(modules),3) 
index.set1<-0
Network=list() 
for (i in 1:length(module.list.set1)){index.set1<-which(c_modules==module.list.set1[i])
Network[[i]]<-row.names(c_modules)[index.set1]}
names(Network)<-module.list.set1
lookup<-function(gene,network){return(network[names(network)[grep(gene,network)]])} 
```

# Filter metadata table and correlate with eigengenes
```{r, fig.height=6, fig.width=5}
nGenes = ncol(datExpr)
nSamples = nrow(datExpr)
MEs = orderMEs(MEs)
ventric$group<-paste0(ventric$trt,"_",ventric$day)
var<-model.matrix(~0+ventric$group)
moduleTraitCor <- cor(MEs, var, use="p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
cor<-melt(moduleTraitCor)
ggplot(cor, aes(Var2, Var1)) + geom_tile(aes(fill = value), 
     colour = "white") + scale_fill_gradient2(midpoint = 0, low = "blue", mid = "white",
                            high = "red", space = "Lab", name="Correlation \nStrength") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + xlab("Treatment") + ylab(NULL)
```

# Get hubgenes in order
```{r}
hubgenes<-lapply(seq_len(length(Network)), function(x) {
  dat<-modulekME[Network[[x]],]
  dat<-dat[order(-dat[paste0("kME",names(Network)[x])]),]
  gene<-rownames(dat)
  return(gene)
})

names(hubgenes)<-names(Network)
d <- unlist(hubgenes)
d <- data.frame(gene = d, 
           vec = names(d))
write_csv(d, path=here("output/glia/wgcna/allglia_wgcna_genemodules_nokme.csv"))
```

# Build linear models for differential expression
```{r, warnings=F, message=F}
MEs %>% select(-MEgrey) -> MEs
data<-data.frame(MEs, day=ventric$day, trt=ventric$trt, 
                 sample=as.factor(ventric$sample), group=ventric$group, 
                 batch=ventric$batch, celltype=Idents(ventric), 
                 groupall=paste0(Idents(ventric), ventric$group))

# loop through modules and perform linear regression
mod<-lapply(colnames(MEs), function(me) {
  # interaction between treatmet group and cell type
  # random effect for batch
  # random effect for sample
  mod<-lmer(MEs[[me]] ~ group*celltype + (1|batch) + (1|sample), data=data)
  # use emmeans to test pariwaise differences
  pairwise<-emmeans(mod, pairwise ~ group|celltype)
  plot<-data.frame(plot(pairwise, plotIt=F)$data)
  sig<-as.data.frame(pairwise$contrasts)
  sig%>%separate(contrast, c("start", "end"), sep = " - ") -> sig
  yvals<-unlist(lapply(unique(sig$celltype), function(x) {
    x<-as.character(x)
    y<-data[data$celltype==x,]
    z<-max(as.numeric(y[[me]]))
    names(z)<-x
    return(z)
  }))
  sig$yvals<-yvals[match(sig$celltype, names(yvals))]
  sig$yvals[duplicated(sig$yvals)]<-sig$yvals[duplicated(sig$yvals)]+.004
  sig$yvals[duplicated(sig$yvals)]<-sig$yvals[duplicated(sig$yvals)]+.004
  sig$yvals[duplicated(sig$yvals)]<-sig$yvals[duplicated(sig$yvals)]+.004
  return(sig)
})

names(mod) <- colnames(MEs)
sig <- bind_rows(mod, .id="id")
sig$symbol <- sig$p.value
# set ranges for significance markings
sig$symbol[findInterval(sig$symbol, c(0.1,2)) == 1L] <-NA
sig$symbol[findInterval(sig$symbol, c(0.01,0.1)) == 1L] <- "*"
sig$symbol[findInterval(sig$symbol, c(0.001,0.01)) == 1L] <- "**"
sig$symbol[findInterval(sig$symbol, c(1e-200,0.001)) == 1L] <- "***" 
data <- melt(data, id.vars = c("day","trt","sample","group","batch","celltype","groupall"))

lapply(unique(data$variable), function(x) {
  tryCatch({
  print(ggplot(data=data[data$variable==x,], aes(x=group, y=as.numeric(value))) + 
  geom_quasirandom(aes(fill=sample), shape=21, size=2, alpha=.75) + 
  scale_fill_manual(values=pal_jco()(10)) +  ylab(NULL) + xlab(NULL) + 
  theme_pubr() + theme(axis.text.x = element_text(angle=45, hjust=1, face="bold"), plot.title = element_text(hjust=0.5)) + 
  scale_y_continuous(aes(name="",limits=c(min(value)-.02,max(value))+.02)) + facet_wrap(.~celltype) +
  labs(y=NULL, x=NULL) + ggtitle(x)) },
  error = function(err) {
    print(err)
  }
  )
})

write_csv(sig, path=here("output/glia/wgcna/allglia_wgcna_linearmodel_testing.csv"))
```

# Plot change at D1 v change at D5 in each model per cell-type
```{r}
sig %>%
  unite(start, end, col = "comparison", remove = F) %>%
  filter(comparison == "FGF_Day-1_PF_Day-1" | comparison == "FGF_Day-5_PF_Day-5") %>%
  unite(estimate, p.value, sep = ",", col = "value") %>%
  dcast(id + celltype ~ start, value.var = "value") %>%
  separate(`FGF_Day-1`, into = c("estimate_1", "p.value_1"), sep = ",") %>%
  separate(`FGF_Day-5`, into = c("estimate_5", "p.value_5"), sep = ",") %>%
  mutate(id = gsub(id, pattern = "ME", replacement = "")) %>%
  mutate(col = if_else(as.numeric(p.value_1) < 0.1 | as.numeric(p.value_5) < 0.1, true = id, false = "white")) %>%
  mutate(sig = if_else(as.numeric(p.value_1) < 0.1 & as.numeric(p.value_5) < 0.1, true = "red",
    false = if_else(as.numeric(p.value_1) < 0.1, true = "blue",
      false = if_else(as.numeric(p.value_5) < 0.1, true = "black", false = "")
    )
  )) %>%
  mutate(id = ifelse(col=="white", yes = NA, no = paste0("sn-glia-M", as.numeric(fct_relevel(as.factor(col), "white", after=Inf))))) -> plot

cols <- unique(plot$col)
names(cols) <- cols

multmod_plot <- ggplot(plot, aes(x = as.numeric(estimate_1), y = as.numeric(estimate_5))) +
  geom_point(data = filter(plot, sig != ""), aes(shape = celltype), size = 4) +
  scale_shape(name="Cell-Type") +
  geom_mark_ellipse(data = plot %>% filter(sig == "blue", p.value_1 != 0.0230780280327769), color = "black", linetype = "dashed") +
  geom_point(size = 5, color = "grey70", aes(shape = celltype)) +
  geom_point(size = 4, aes(color = factor(id), shape = celltype)) + 
  coord_flip(clip="off") + geom_hline(yintercept = 0, linetype="dashed") +
  scale_color_npg(name = "Modules", na.value="grey",
                  labels=c("sn-glia-M1","sn-glia-M2","sn-glia-M3","sn-glia-M4","sn-glia-M5","p>0.05")) +
  geom_vline(xintercept = 0, linetype="dashed") +
  annotate(geom = "curve", x = 0.045, y = 0.02, xend = 0.03, yend = 0.005, curvature = .3, 
           arrow = arrow(length = unit(2, "mm"))) + 
  annotate(geom = "label", x = 0.045, y = 0.0225, label = "Up at D1", hjust = "left", ) +
  annotate(geom = "curve", x = -0.01, y = 0.03, xend = -0.025, yend = 0.035,curvature = .3, 
           arrow = arrow(length = unit(2, "mm"))) + 
  annotate(geom = "label", x = -0.005, y = 0.025, label = "Down at D1\n Up at D5", hjust = "left") +
  annotate(geom = "curve", x = 0.01, y = 0.015, xend = -0.011, yend = 0.02, curvature = .3, 
           arrow = arrow(length = unit(2, "mm"))) + 
  annotate(geom = "label", x = 0.015, y = 0.01, label = "Up at D5", hjust = "left") +  
  annotate(geom = "curve", x = -0.02, y = -0.009, xend = -0.013, yend = 0.002, curvature = .3, 
           arrow = arrow(length = unit(2, "mm"))) + 
  annotate(geom = "label", x = -0.02, y = -0.02, label = "Down at D1", hjust = "left") + 
  theme_bw() + 
  xlab(expression(bold(Day1~beta))) +
  ylab(expression(bold(Day5~beta))) + 
  guides(shape = guide_legend(override.aes = list(size=4), 
                              title.theme = element_text(face="bold", size=8),
                              label.theme = element_text(face="bold", size=8)),
         color =  guide_legend(title.theme = element_text(size=8, face="bold"), 
                               label.theme = element_text(face="bold", size=8))) +
  theme_figure

ggsave(here("data/figures/wgcna_res.png"), w=8, h=5)
```

# Plot gene networks
```{r, fig.width=12, fig.height=5}
hubgenes <- lapply(seq_len(length(Network)), function(x) {
  dat <- modulekME[Network[[x]], ]
  dat <- dat[order(-dat[paste0("kME", names(Network)[x])]), ]
  gene <- data.frame(gene=rownames(dat),kme=dat[,x]) 
  return(gene)
})

names(hubgenes)<- names(Network)

color <- c("black","green","magenta")
lapply(color, function(col) {
  maxsize <- 15
  hubs <- data.frame(genes=hubgenes[[col]]$gene[1:maxsize], kme = hubgenes[[col]]$kme[1:maxsize], mod = rep(col,15))
}) -> hub_plot 


hub_plot <- lapply(hub_plot, function(x) {
  adj[as.character(x$genes), as.character(x$genes)] %>% 
  graph.adjacency(mode = "undirected", weighted = T, diag = FALSE) %>%
  as_tbl_graph(g1) %>% upgrade_graph() %>% activate(nodes) %>%
    dplyr::mutate(mod=x$mod) %>%
    dplyr::mutate(kme=x$kme)  %>% 
    activate(edges) %>% dplyr::filter(weight>.15)}
)

hub_plot <- lapply(hub_plot, function(x) {
  x  %>% 
    activate(nodes) %>% 
    dplyr::mutate(color = ifelse(name %in% c("Gfap","Itgav","Vim","Sprr1a"), yes="red", no="black"))
})



set.seed("139")
plotlist <- lapply(hub_plot,  function(x) {
  print(ggraph(x,  layout = 'fr') + 
  geom_edge_link(color="darkgrey", aes(alpha = weight), show.legend = F)  + 
  scale_edge_width(range = c(0.2, 1)) + geom_node_text(aes(label = name, color=color), fontface="bold", size=3, repel = T) +
  geom_node_point(shape=21, alpha=0.5, fill="grey70", size=3) + 
    scale_color_manual(values=c("gray0","red")) +
  theme_graph() + theme(legend.position = "none", plot.title = element_text(hjust=0.5, vjust=1), plot.margin = unit(c(0, 0, 0, 0), "cm")) + coord_cartesian(clip = "off"))
})

genenet <- plot_grid(plotlist[[1]], plotlist[[2]], plotlist[[3]], ncol=1, labels=c("f"), scale=0.9)
```
```{r, fig.width=12, fig.height=12}
load(here("data/figures/fig3/fig3_top.RData"))
fig_bot <- plot_grid(multmod_plot, genenet, nrow=1, align="hv", axis="tblr", labels=c("e",""), rel_widths =c(1.25,1), scale=c(0.9,1))
save(fig_bot, file = here("data/figures/fig3/fig3_bot.RData"))
fig_full <- plot_grid(fig3_top,fig_bot, ncol=1, rel_heights = c(1.5,1))
ggsave2(fig_full, filename = here("data/figures/fig3/fig3_arranged.tiff"), h=12, w=12, dpi = 600, compression="lzw")
fig_full
```


# Find GO term enrichment of modules
```{r, fig.height=10}
go_col <- unique(plot$col)[-2]
goterms<-lapply(go_col, function(x) { 
  gprofiler(as.character(hubgenes[[x]]$gene), ordered_query = T, 
               organism = "mmusculus", significant = T, custom_bg = colnames(datExpr),
                           src_filter = c("GO:BP","REAC","KEGG"), hier_filtering = "strong",
                           min_isect_size = 3, 
                           sort_by_structure = T,exclude_iea = T, 
                           min_set_size = 10, max_set_size = 300,correction_method = "fdr") %>% 
    arrange(p.value) -> got
  return(got)
})


names(goterms) <- go_col
goterms %>% bind_rows(.id="id") %>%
  mutate(padj=p.adjust(p.value, "fdr")) -> godat

write_csv(godat, path=here("output/glia/wgcna/allglia_wgcna_goterms.csv"))

godat %>% group_by(id) %>% filter(id %in% c("black","green")) %>% arrange(p.value) %>% slice(1:5) %>% 
  select(p.value, padj, term.name, domain, id) %>% arrange(id) %>%
  ggplot(aes(x=str_wrap(term.name,20), y=-log10(padj), fill=domain)) + geom_col() +
  scale_fill_npg() +
  facet_wrap(.~id, scales="free_y", ncol=1) + theme_pubr() + 
  theme(text = element_text(size=7), 
        axis.text.x = element_text(angle=45, hjust=1)) + coord_flip() +
  xlab("GO Term") + geom_hline(yintercept = -log10(0.05), linetype="dashed", size=1) + 
  labs_pubr()

ggsave(filename = here("output/glia/wgcna/allglia_goterm.png"), h=6, w=8)
```

